\section{Professional Experience}

\textbf{Senior Data Engineer, Platform \& Data Engineering} \hfill Dec 2023 -- Present\\
\textit{Ennoble First \textbar{} SAFFIRE / CHINOOK}

\begin{itemize}
    \item Define and advance enterprise data platform capabilities supporting distributed processing, governed data access, and scalable analytics across engineering and mission teams.
    \item Establish data engineering standards for ingestion, transformation, validation, and lifecycle management, improving reliability and reducing organizational rework across teams.
    \item Lead evolution of medallion-based architecture patterns, enabling consistent data lifecycle management and scalable consumption across multiple teams.
    \item Drive cross-functional alignment across engineering, architecture, operations, and leadership, translating requirements into scalable platform capabilities supporting mission and organizational objectives.
    \item Improve accessibility, consistency, and trust in enterprise data, enabling more reliable and timely decision-making across engineering and mission stakeholders.
    \item Influence proposal and capture efforts by shaping technical solution approaches and validating feasibility aligned to mission and business requirements.
    \item Contribute to development of reusable architectural patterns and technical frameworks that guide data platform modernization and distributed system design across the organization.
\end{itemize}

\vspace{4pt}

\textbf{Software Engineer / Data Engineer} \hfill May 2021 -- Dec 2023\\
\textit{Army Software Factory \textbar{} Army Futures Command}

\begin{itemize}
    \item Led architecture and scaling of high-volume data pipelines processing 500M+ records, reducing processing time from 24 hours to 26 minutes and enabling near real-time, decision-critical data availability.
    \item Delivered curated, enterprise-grade datasets into mission-critical platforms (Army Vantage), improving data accessibility and operational visibility across engineering and leadership stakeholders.
    \item Drove modernization of fragmented data systems into unified, scalable architectures, increasing interoperability and enabling enterprise-wide data access.
    \item Partnered across engineering, operations, and product stakeholders to align data platform capabilities with mission objectives and long-term organizational priorities.
    \item Enabled adoption of scalable data engineering practices, improving reliability, consistency, and delivery velocity across teams.
    \item Delivered curated, enterprise-grade datasets into mission-critical platforms, improving data accessibility, operational visibility, and speed of decision-making across stakeholders.
\end{itemize}

\vspace{4pt}

\textbf{Data Engineer} \hfill Apr 2021 -- Sep 2022\\
\textit{Texas ProTax}

\begin{itemize}
    \item Designed and implemented data integration workflows across SQL and NoSQL systems, establishing scalable pipelines to support enterprise analytics and machine learning initiatives.
    \item Delivered analysis-ready datasets aligned with governance and modeling standards, improving data usability and accelerating downstream decision-making.
    \item Supported executive-level data initiatives through alignment of data architecture with business and operational priorities.
\end{itemize}

\vspace{4pt}

\textbf{Digital Intelligence Systems Master Gunner} \hfill Mar 2017 -- Apr 2021\\
\textit{United States Army}

\begin{itemize}
    \item Led integration and modernization of enterprise intelligence data platforms (DCGS-A CD2, Palantir-powered), aligning data architecture with mission-critical decision systems across operational units.
    \item Directed and coordinated cross-functional teams spanning analysts, engineers, and operational personnel, ensuring effective execution of data-driven mission objectives.
    \item Partnered with leadership and technical stakeholders to translate complex operational requirements into scalable data platform capabilities.
    \item Owned reliability, accuracy, and availability of mission-critical data systems, enabling consistent real-time decision support in high-pressure environments.
    \item Drove adoption of data governance standards and structured data practices, improving consistency and effectiveness across distributed teams and organizations.
\end{itemize}

\vspace{4pt}

\textbf{Intelligence Analyst $\rightarrow$ Data Scientist $\rightarrow$ Data Architect} \hfill Sep 2002 -- Apr 2024\\
\textit{United States Army}

\begin{itemize}
    \item Progressed through increasingly senior roles with responsibility for leading data-driven operations and aligning data capabilities with mission-critical decision-making across multiple organizational levels.
    \item Led and mentored personnel from enlisted through warrant officer ranks, enabling career progression, leadership development, and technical specialization across operational units.
    \item Developed and enabled cross-functional workforce capability by introducing technical training pathways, including industry-recognized certifications (e.g., CompTIA), expanding analytical and engineering skill sets across teams.
    \item Supported partner nation information sharing and interoperability efforts, mentoring international personnel and improving technical competency and collaboration across joint and multinational environments.
    \item Contributed to doctrine development and analyst training standards at the U.S. Army Intelligence Center of Excellence, influencing how data and intelligence workflows are executed across the force.
    \item Designed scalable data architectures and standardized analytical workflows, enabling consistent, repeatable, and reliable decision-support across distributed operational environments.
    \item Bridged analytics, engineering, and leadership stakeholders, translating operational needs into scalable data solutions aligned to organizational and mission objectives.
\end{itemize}